\subsection*{Declaration of acceptance}

This thesis fulfils all formal and content requirements prescribed by the Faculty of Mechanical Engineering of Budapest University of Technology and Economics, as well as it fully complies tasks specified in the transcript. I consider this thesis as it is suitable for submission for public review and for public presentation.

Done at Budapest, \today

\begin{flushright}
\vspace*{1cm}
\makebox[7cm]{\rule{6cm}{.4pt}}\\
\makebox[7cm]{\emph{Dr. Dénes Takács}}\\
\end{flushright}

\vspace*{1.5cm}

\subsection*{Student declaration}

I, Gergely Németh, the undersigned, hereby declare that the present thesis work has been prepared by myself without any unauthorized help or assistance such that only the specified sources (references, tools, etc.) were used. All parts taken from other sources word by word or after rephrasing but with identical meaning were unambiguously identified with explicit reference to the sources utilized.

I authorize the Faculty of Mechanical Engineering of the Budapest University of Technology and Economics to publish the principal data of the thesis work (author's name, title, abstracts in English and in Hungarian, year of preparation, supervisor's name, etc.) in a searchable, public, electronic and online database and to publish the full text of the thesis work on the internal network of the university (or for authenticated users). I declare that the submitted hard copy of the thesis work and its electronic version are identical. Full text of thesis works classified upon the decision of the Dean will be published after two years.

Done at Budapest, \today

\begin{flushright}
\vspace*{1cm}
\makebox[7cm]{\rule{6cm}{.4pt}}\\
\makebox[7cm]{\emph{Gergely Németh}}\\
\end{flushright}